Compositional Equivalences Based on Open pNets

### 2022-09-28 2016-Weak-FH-Bisim/WeakBisim.tex

Another aspect of the logical interpretation of the
formulas is that we make no distinction between the equality and the
equivalence on boolean formulas, i.e. equivalence of two predicates
$\Pred$ and $\Pred\,'$ can be denoted $\Pred=\Pred\,'$, where the $=$ symbol is not interpreted in a syntactical way.

@@

The equivalence we need is a strong bisimulation between
open automata having exactly the same holes (same indices and same sorts), but using a
flexible matching 
between open transitions, this will allow us to compare pNets
with different architectures.

@@

One of the original aspects of FH-bisimulation is due to the symbolic nature of open automata. Indeed, a single state of the automaton represents a potentially infinite number of concrete states, depending on the value of the automaton variables, and a single open transition of the automaton may also be instantiated with an unbounded number of values for the transition parameters. Consequently it would be too restrictive to impose that each transition of one automaton is matched by exactly one transition of the bisimilar automaton. Thus the definition of bisimulation requires that, for each open transition of one automaton, there exists a matching  set of open transitions covering the original one, indeed depending on the value of action parameters or automaton variables, different open transitions might simulate the same one.

The parametric nature of the automata entails a second original aspect of FH-bisimulation: the nature of the bisimulation relation itself.
 A classical relation between states can be seen as a function mapping pairs of state to a boolean value (true if the states are related, false if they are not). An FH-bisimulation relation maps pairs of states to boolean expressions that use variables of the two systems.  %Each variable used in $\Pred_{s,t}$ should refer to a state variable of $s$ or $t$ and there must be no ambiguity on the referenced variable, we thus require that the two compared automata have disjoint state variables.
Formally, a relation over the states of two open automata  $\mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $\mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$ has the signature $\mathcal{S}_1\times\mathcal{S}_2\to\AlgB$. 
We suppose without loss of generality that the variables of the two open automata are disjoint.
We adopt a notation similar to standard relations and denote it
 $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$, where: 1) For any pair $(s,t)\in \mathcal{S}_1\times \mathcal{S}_2$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$       is 
      True, i.e. the states are related when the value of the automata variables% in $s$ and $t$
  verify the predicate $\Pred_{s,t}$. 2) The free variables of  $\Pred_{s,t}$ belong to $V_1$ and $V_2$, i.e. $\vars(\Pred_{s,t})\subseteq V_1\cup V_2$.
 FH-bisimulation is defined formally\footnote{In this article, we denote $\beta_{j x}$  a double indexed set, instead of the classical $\beta_{j,\, x}$. Indeed the standard notation would be too heavy in our case.}: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim} ~\\
\noindent
	Suppose 
   $A_1 = \mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $A_2 = \mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$
   are open automata with identical holes of the same sort, with disjoint sets of variables ($V_1\cap V_2=\emptyset$).

 Then 
$\mathcal{R}$ is an FH-bisimulation if and only if for any  states
$s\in\mathcal{S}_1$ and $t\in\mathcal{S}_2$, $(s,t|\Pred_{s,t})\in\mathcal{R}$, we 
have
the following:

>>>>

Another aspect of the semantic interpretation is that we consider terms up to semantic equivalence, i.e. equivalence of two predicates
$\Pred$ and $\Pred\,'$ can be denoted $\Pred=\Pred\,'$, where the $=$ symbol is interpreted semantically.

@@

%
%The equivalence we need is a strong bisimulation between
%open automata having exactly the same holes (same indices and same sorts), but using a
%flexible matching 
%between open transitions, this will allow us to compare pNets
%with different architectures.

@@

One of the original aspects of FH-bisimulation is due to the symbolic nature of open automata. Indeed, a single state of the automaton represents a potentially infinite number of concrete states, depending on the value of the automaton variables, and a single open transition of the automaton may also be instantiated with an unbounded number of values for the transition parameters. Consequently it would be too restrictive to impose that each transition of one automaton is matched by exactly one transition of the bisimilar automaton. Thus the definition of bisimulation requires that, for each open transition of one automaton, there exists a matching  set of open transitions covering the original one. Indeed depending on the value of action parameters or automaton variables, different open transitions might simulate the same one.

The parametric nature of the automata entails a second original aspect of FH-bisimulation: the nature of the bisimulation relation itself.
 A classical relation between states can be seen as a function mapping pairs of state to a boolean value (true if the states are related, false if they are not). An FH-bisimulation relation maps pairs of states to boolean expressions that use variables of the two systems.  %Each variable used in $\Pred_{s,t}$ should refer to a state variable of $s$ or $t$ and there must be no ambiguity on the referenced variable, we thus require that the two compared automata have disjoint state variables.
Formally, a relation over the states of two open automata  $\mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $\mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$ has the signature $\mathcal{S}_1\times\mathcal{S}_2\to\AlgB$. 
We suppose without loss of generality that the variables of the two open automata are disjoint.
We adopt a notation similar to standard relations and denote it
 $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$, where: 1) For any pair $(s,t)\in \mathcal{S}_1\times \mathcal{S}_2$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$       is 
      True, i.e. the states are related when the value of the automata variables % in $s$ and $t$
  satisfy the predicate $\Pred_{s,t}$. 2) The free variables of  $\Pred_{s,t}$ belong to $V_1$ and $V_2$, i.e. $\vars(\Pred_{s,t})\subseteq V_1\cup V_2$.
 FH-bisimulation is defined formally\footnote{In this article, we denote $\beta_{j x}$  a double indexed set, instead of the classical $\beta_{j,\, x}$. Indeed the standard notation would be too heavy in our case.}: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim} ~\\
\noindent
	Suppose 
   $A_1 = \mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $A_2 = \mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$
   are open automata with identical holes of the same sort, with disjoint sets of variables ($V_1\cap V_2=\emptyset$).

 Then 
$\mathcal{R}$ is an FH-bisimulation if and only if for all  states
$s\in\mathcal{S}_1$ and $t\in\mathcal{S}_2$, $(s,t|\Pred_{s,t})\in\mathcal{R}$, we 
have
the following:
